---
title: "Create mater table for nanopore WTC-11 data ENCFF961HLO"
author: "Vladimir Souza"
date: "5/12/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Introduction

Create the master table for WTC-11 data.  

Compare: DeepVariant and Clair3 combined with SNCR, flagCorrection or alone, and GATK with SNCR.  

Nanopore RNA-seq raw sequencing data:  
  * only primary alignments (secondary and supplementary alignments were filtered out);  
  * duplicates not marked/removed.  

sort-read:  
  * VCF downloaded from Allen Institute;  
  * only chromosomes chr1, chr2, ..., chr22, chrX, chrY;  
  * intronic regions removed.  




# Input variables

```{r}
### methods to validate
# name of the methods to validate
METHOD_NAMES <- c("c3", "c3_sncr", "c3_sncr_fc", "gatk",
                  "dv", "dv_sncr", "dv_sncr_fc")
# name of the dataset used with the methods to validate
METHOD_DATASET_NAME <- "Nanopore"
# VCF files
METHOD_VCF_FILES <- c(
  "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/files/ENCFF961HLO/variant_calling_from_nanopore/clair3/c3/pileup_pass.vcf.gz",
  "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/files/ENCFF961HLO/variant_calling_from_nanopore/clair3/c3_sncr/pileup_pass.vcf.gz",
  "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/files/ENCFF961HLO/variant_calling_from_nanopore/clair3/c3_sncr_fc/pileup_pass.vcf.gz",
  "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/files/ENCFF961HLO/variant_calling_from_nanopore/gatk/nanopore_wtc11.recal_pass.vcf.gz",
  "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/files/ENCFF961HLO/variant_calling_from_nanopore/deepvariant/dv/deepvariant_calls_pass.vcf.gz",
  "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/files/ENCFF961HLO/variant_calling_from_nanopore/deepvariant/dv_sncr/deepvariant_calls_pass.vcf.gz",
  "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/files/ENCFF961HLO/variant_calling_from_nanopore/deepvariant/dv_sncr_fc/deepvariant_calls_pass.vcf.gz"
)
# BAM of the data
METHOD_BAM_FILE <- "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/files/ENCFF961HLO/data_manipulation/aln_s.bam"
### ground-truth
# name
TRUTH_NAME <- "allen"
# name of the dataset used to generate the ground-truth
TRUTH_DATASET_NAME <- "shortRead"
# VCF file
TRUTH_VCF_FILE <- "/home/vbarbo/project_2021/paper_analysis/wtc11/ground_truth/3546dc62_AH77TTBBXX_DS-229105_GCCAAT_recalibrated_subsetChromosomes_pass.vcf.gz"
# BAM file
# downloaded by script /home/vbarbo/project_2021/projects/lrRNAseqVariantCalling/1_download_data/wtc11/not_filtered_ground_truth_vcf/wtc11_vcf.sh
TRUTH_BAM_FILE <- "/home/vbarbo/project_2021/paper_analysis/wtc11/ground_truth/84773251_trimmed.AH77TTBBXX_DS-229105_GCCAAT.sorted.rg.final.1.bam"
### variables
MAX_DIST_FROM_SPLICE_SITE <- 20
THREADS <- 30
GENOME_REF_FILE <- "/home/vbarbo/project_2021/paper_analysis/reference/genome/GRCh38.p13_all_chr.fasta"
```




# libraries

```{r}
library(lrRNAseqBenchmark)
library(GenomicAlignments)
library(dplyr)
library(sarlacc)
library(snakecase)
```




# Create the master table

## Initiate the master table

Get variant positions and which method could call them.

```{r}
dat <- initiate_master_table(
  METHOD_VCF_FILES[1],
  METHOD_VCF_FILES[2],
  METHOD_VCF_FILES[3],
  METHOD_VCF_FILES[4],
  METHOD_VCF_FILES[5],
  METHOD_VCF_FILES[6],
  METHOD_VCF_FILES[7],
  TRUTH_VCF_FILE,
  method_names=c(METHOD_NAMES, TRUTH_NAME)
)
```


## Add QD tag from ground-truth VCF

```{r}
# dat <- add_info_tag_from_vcf(dat, "qd_allen", "QD", TRUTH_VCF_FILE)
```


## Add dp tag for the ground truth.

```{r}
# dat <- add_info_tag_from_vcf(dat, "dp_info_allen", "DP", TRUTH_VCF_FILE)
```


## Add QUAL tag from VCF

```{r}
method_names_tlc <- to_lower_camel_case(METHOD_NAMES)
dat <- add_qual_from_vcf(dat, method_names_tlc[1], METHOD_VCF_FILES[1])
dat <- add_qual_from_vcf(dat, method_names_tlc[2], METHOD_VCF_FILES[2])
dat <- add_qual_from_vcf(dat, method_names_tlc[3], METHOD_VCF_FILES[3])
dat <- add_qual_from_vcf(dat, method_names_tlc[4], METHOD_VCF_FILES[4])
dat <- add_qual_from_vcf(dat, method_names_tlc[5], METHOD_VCF_FILES[5])
dat <- add_qual_from_vcf(dat, method_names_tlc[6], METHOD_VCF_FILES[6])
dat <- add_qual_from_vcf(dat, method_names_tlc[7], METHOD_VCF_FILES[7])
```


## get splice site positions from the BAM file and add columns about splice sites

```{r}
method_bam <- readGAlignments(METHOD_BAM_FILE)
# splice_sites <- get_splice_sites_info(method_bam, THREADS)

# saveRDS(splice_sites, file="/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/extra_files/splice_sites_nanopore_wtc11_ENCFF961HLO.rds")
splice_sites <- readRDS("/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/extra_files/ENCFF961HLO/splice_sites_nanopore_wtc11_ENCFF961HLO.rds")

dat <- add_splice_site_info_to_master_table(
  dat,
  splice_sites,
  MAX_DIST_FROM_SPLICE_SITE
)
```


## Add the read coverage (from the BAM file) of each variant

Add iso-seq read coverage.

```{r}
method_coverage <- coverage(METHOD_BAM_FILE)
dat <- add_read_coverage_from_bam_to_master_table(
  dat,
  method_coverage,
  METHOD_DATASET_NAME
)
```

Add short-read coverage (ground truth).

```{r}
### get the coverage of the ground-truth short-read bam.
### for this, we load the objects created in script 'lrRNAseqVariantCalling/isoseq_variant_calling_analysis_code/4_create_master_table/wtc11/mt_wtc11_allMethods_v7.Rmd
### The object `truth_coverage_1` stores the coverage of chromosomes 1--12, and `truth_coverage_2` of 13--22 + chrX + chrY.
truth_coverage_1 <- readRDS("/home/vbarbo/project_2021/paper_analysis/extra_files/truth_coverage_1_wtc11.rds")
truth_coverage_2 <- readRDS("/home/vbarbo/project_2021/paper_analysis/extra_files/truth_coverage_2_wtc11.rds")


### add short-read coverage to master table. must do it in two steps: first the chromosomes in truth_coverage_1 and lately the ones in truth_coverage_2
dat_1 <- filter(dat, chrm %in% names(truth_coverage_1))
dat_2 <- filter(dat, chrm %in% names(truth_coverage_2))
dat_1 <- add_read_coverage_from_bam_to_master_table(
  dat_1,
  truth_coverage_1,
  TRUTH_DATASET_NAME
)
dat_2 <- add_read_coverage_from_bam_to_master_table(
  dat_2,
  truth_coverage_2,
  TRUTH_DATASET_NAME
)
dat <- rbind(dat_1, dat_2) %>% 
  arrange(chrm, pos)
```


## Add the number of N-cigar reads per site

```{r}
dat <- add_number_of_n_cigar_reads_to_master_table(
  dat, 
  method_bam, 
  METHOD_DATASET_NAME
)
```


## Add column to classify the method calls (compare to the ground-truth)

```{r}
### method 1
dat <- add_method_vs_truth_comparison_to_master_table(
  dat,
  METHOD_NAMES[1],
  TRUTH_NAME
)
### method 2
dat <- add_method_vs_truth_comparison_to_master_table(
  dat,
  METHOD_NAMES[2],
  TRUTH_NAME
)
### method 3
dat <- add_method_vs_truth_comparison_to_master_table(
  dat,
  METHOD_NAMES[3],
  TRUTH_NAME
)
### method 4
dat <- add_method_vs_truth_comparison_to_master_table(
  dat,
  METHOD_NAMES[4],
  TRUTH_NAME
)
### method 5
dat <- add_method_vs_truth_comparison_to_master_table(
  dat,
  METHOD_NAMES[5],
  TRUTH_NAME
)
### method 6
dat <- add_method_vs_truth_comparison_to_master_table(
  dat,
  METHOD_NAMES[6],
  TRUTH_NAME
)
### method 7
dat <- add_method_vs_truth_comparison_to_master_table(
  dat,
  METHOD_NAMES[7],
  TRUTH_NAME
)
```


## Add columns to inform the variant type according each method and the ground truth

For the ground-truth

```{r}
### ground truth
dat <- add_variant_type_to_master_table(dat, TRUTH_VCF_FILE, TRUTH_NAME)
### methods
dat <- add_variant_type_to_master_table(dat, METHOD_VCF_FILES[1],
                                        METHOD_NAMES[1])
dat <- add_variant_type_to_master_table(dat, METHOD_VCF_FILES[2],
                                        METHOD_NAMES[2])
dat <- add_variant_type_to_master_table(dat, METHOD_VCF_FILES[3],
                                        METHOD_NAMES[3])
dat <- add_variant_type_to_master_table(dat, METHOD_VCF_FILES[4],
                                        METHOD_NAMES[4])
dat <- add_variant_type_to_master_table(dat, METHOD_VCF_FILES[5],
                                        METHOD_NAMES[5])
dat <- add_variant_type_to_master_table(dat, METHOD_VCF_FILES[6],
                                        METHOD_NAMES[6])
dat <- add_variant_type_to_master_table(dat, METHOD_VCF_FILES[7],
                                        METHOD_NAMES[7])
```


## Add columns to inform whether the variant is a indel or not (-1, 0, 1, or NA)

* 1 = is indel
* 0 = is not indel
* -1 = non defined (heterozygous alternative) (truth's gt has priority over the method's)
* NA = variant not calls by the method or the truth

```{r}
### truth
dat <- is_indel_method(dat, TRUTH_NAME)

### methods
dat <- is_indel_method(dat, TRUTH_NAME, METHOD_NAMES[1])
dat <- is_indel_method(dat, TRUTH_NAME, METHOD_NAMES[2])
dat <- is_indel_method(dat, TRUTH_NAME, METHOD_NAMES[3])
dat <- is_indel_method(dat, TRUTH_NAME, METHOD_NAMES[4])
dat <- is_indel_method(dat, TRUTH_NAME, METHOD_NAMES[5])
dat <- is_indel_method(dat, TRUTH_NAME, METHOD_NAMES[6])
dat <- is_indel_method(dat, TRUTH_NAME, METHOD_NAMES[7])
```


## Add variant/sequencing_error dencity per variant

```{r}
### TRUTH_NAME
dat <- add_variant_density_of_a_method(dat, 201, TRUTH_NAME)
### all methods
dat <- add_variant_density_of_a_method(dat, 201, METHOD_NAMES)

dat <- add_variant_density_of_a_method(dat, 201, METHOD_NAMES[c(2,3,7)] )
dat <- add_variant_density_of_a_method(dat, 201, METHOD_NAMES[c(2,3,5)] )
dat <- add_variant_density_of_a_method(dat, 201, METHOD_NAMES[c(2,3,4,7)] )
dat <- add_variant_density_of_a_method(dat, 201, METHOD_NAMES[c(2,3,4,5)] )
```









```{r}
dat_beforeAddingGenotypes <- dat
saveRDS(dat_beforeAddingGenotypes,
        file="~/keita/dat_beforeAddingGenotypes.r")
```










<!-- dat_bk2 <- dat -->
<!-- dat <- dat_bk2 -->

## add genotype

```{r}
dat <- add_format_tag_from_vcf(dat, TRUTH_NAME, "GT", TRUTH_VCF_FILE)
dat <- add_format_tag_from_vcf(dat, METHOD_NAMES[1], "GT", METHOD_VCF_FILES[1])
dat <- add_format_tag_from_vcf(dat, METHOD_NAMES[2], "GT", METHOD_VCF_FILES[2])
dat <- add_format_tag_from_vcf(dat, METHOD_NAMES[3], "GT", METHOD_VCF_FILES[3])
dat <- add_format_tag_from_vcf(dat, METHOD_NAMES[4], "GT", METHOD_VCF_FILES[4])
dat <- add_format_tag_from_vcf(dat, METHOD_NAMES[5], "GT", METHOD_VCF_FILES[5])
dat <- add_format_tag_from_vcf(dat, METHOD_NAMES[6], "GT", METHOD_VCF_FILES[6])
dat <- add_format_tag_from_vcf(dat, METHOD_NAMES[7], "GT", METHOD_VCF_FILES[7])
# sel_cols <- paste0( "gt_", c(TRUTH_NAME, METHOD_NAMES[c(3,7,8,9)]) )
sel_cols <- paste0( "gt_", c(TRUTH_NAME, METHOD_NAMES) )
k <- is.na( dat[,sel_cols] )
dat[,sel_cols] [k] <- "0/0"
```













## add homopolymer length (from reference FASTA), but only for 0/1 and 1/1 genotypes and indels

```{r}
# # get all homopolymers of the reference genome
# genome_ref <- readDNAStringSet(GENOME_REF_FILE)
# homopolymers <- homopolymerFinder(genome_ref)
# names(homopolymers) <- sub("(^chr[0-9]+|X|Y).*", "\\1", names(homopolymers))
### load file with all homopolymers of the reference genome created in script /home/vbarbo/project_2021/projects/lrRNAseqVariantCalling/4_create_master_table/jurkat/mt_jurkat_allMethods_v7.Rmd
homopolymers <- readRDS("/home/vbarbo/project_2021/paper_analysis/extra_files/homopolymers_GRCh38.p13_all_chr.rds")
dat <- add_homopolymer_length_when_indels(dat, homopolymers)
```


## Take a look at the master table

```{r}
head(dat)
dat_bk <- dat
```


## Save master table to a file

```{r}
mt_wtc11_clair_gatk_dv_notFiltered <- dat
saveRDS(mt_wtc11_clair_gatk_dv_notFiltered, "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/extra_files/ENCFF961HLO/clair3_gatk_deepvariant/master_tables/mt_wtc11_clair_gatk_dv_notFiltered_v1.rds")
```




# Filtering master table

Sites to keep:  
* short-read variant density: x <= 3                (ignore positions)  
* ignore intronic regions (isoSeq_coverage > 0)     (ignore positions)  
* DeepVariant QUAL: x >= 15                         (remove method calling)  
* short-read coverare : quantil 5% <= x <= 95%      (ignore positions)  

```{r}
# no filtering
dat1 <- dat
dim(dat1)

k <- paste0(TRUTH_DATASET_NAME, "_coverage")
truth_cover_quantiles <- quantile(dat1[,k], probs=c(.05, .95))

# variant/sequence_error density -- filter using the ground truth
k <- paste0("variantDensity_", TRUTH_NAME)
k <- !( dat1[,k] > 3 )
dat1 <- dat1[k,]
dim(dat1)

# ignore intronic regions (iso-seq coverage equal to zero)
k <- paste0(METHOD_DATASET_NAME, "_coverage")
k <- dat1[,k] > 0
dat1 <- dat1[k,]
dim(dat1)

# ignore regions based on short-read (ground-truth) coverage quantiles: keep 5% <= x <= 95%
k <- paste0(TRUTH_DATASET_NAME, "_coverage")
k <- dat1[,k] >= truth_cover_quantiles[1] &
  dat1[,k] <= truth_cover_quantiles[2]
dat1 <- dat1[k,]
dim(dat1)

# # Filter out DeepVariant calls by QUAL -- but keep the position
# # dv
# k <- which( dat1$qual_dv < 10 )
# dat1$in_dv[k] <- 0
# dat1 <- add_method_vs_truth_comparison_to_master_table(
#   dat1,
#   METHOD_NAMES[5],
#   TRUTH_NAME,
#   replace_column=TRUE
# )
# # dv_sncr
# k <- which( dat1$qual_dvSncr < 10 )
# dat1$in_dv_sncr[k] <- 0
# dat1 <- add_method_vs_truth_comparison_to_master_table(
#   dat1,
#   METHOD_NAMES[6],
#   TRUTH_NAME,
#   replace_column=TRUE
# )
# # dv_sncr_fc
# k <- which( dat1$qual_dvSncrFc < 10 )
# dat1$in_dv_sncr_fc[k] <- 0
# dat1 <- add_method_vs_truth_comparison_to_master_table(
#   dat1,
#   METHOD_NAMES[7],
#   TRUTH_NAME,
#   replace_column=TRUE
# )

### don't filter by DeepVariant QUAL

### don't filter by Clair3 QUAL
```

save filtered master table

```{r}
mt_wtc11_clair_gatk_dv_filtered <- dat1
saveRDS(mt_wtc11_clair_gatk_dv_filtered, "/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/extra_files/ENCFF961HLO/clair3_gatk_deepvariant/master_tables/mt_wtc11_clair_gatk_dv_filtered_v1.rds")
```



### see clair3 QUAL distribution for TPs and FPs

```{r}
dat1 <- readRDS("/home/vbarbo/project_2021/paper_analysis/secondary_analyses/nanopore/extra_files/master_tables/mt_wtc11_clairPileupMerge_filtered_v1.rds")

library(ggplot2)

png("~/distribution_qual_nanopore_ENCFF961HLO_wtc11_c3.png", width=800)
ggplot(dat1, aes(x=qual_c3, fill=c3_classification)) +
  geom_histogram(alpha=.5, binwidth=1, position = 'identity') +
  geom_vline(xintercept=c(5,10,15))
dev.off()

png("~/distribution_qual_nanopore_ENCFF961HLO_wtc11_c3Sncr.png", width=800)
ggplot(dat1, aes(x=qual_c3Sncr, fill=c3_sncr_classification)) +
  geom_histogram(alpha=.5, binwidth=1, position = 'identity') +
  geom_vline(xintercept=c(5,10,15))
dev.off()

png("~/distribution_qual_nanopore_ENCFF961HLO_wtc11_c3SncrFc.png", width=800)
ggplot(dat1, aes(x=qual_c3SncrFc, fill=c3_sncr_fc_classification)) +
  geom_histogram(alpha=.5, binwidth=1, position = 'identity') +
  geom_vline(xintercept=c(5,10,15))
dev.off()

############## we can filter by qual when using p-m-dv
```




### see DeepVariant QUAL distribution for TPs and FPs (Iso-Seq data)

```{r}
dat1 <- readRDS("....../home/vbarbo/project_2021/paper_analysis/extra_files/master_tables/mt_wtc11_allMethods_notFiltered_v7.rds")

library(ggplot2)

png("~/distribution_qual_nanopore_ENCFF961HLO_wtc11_dv_wesModel.png", width=800)
ggplot(dat1, aes(x=qual_dv, fill=dv_classification)) +
  geom_histogram(alpha=.5, binwidth=1, position = 'identity') +
  geom_vline(xintercept=c(5,10,15))
dev.off()

png("~/distribution_qual_nanopore_ENCFF961HLO_wtc11_dvS_wesModel.png", width=800)
ggplot(dat1, aes(x=qual_dvSncr, fill=dv_sncr_classification)) +
  geom_histogram(alpha=.5, binwidth=1, position = 'identity') +
  geom_vline(xintercept=c(5,10,15))
dev.off()

png("~/distribution_qual_nanopore_ENCFF961HLO_wtc11_dvSFc_wesModel.png", width=800)
ggplot(dat1, aes(x=qual_dvSncrFc, fill=dv_sncr_fc_classification)) +
  geom_histogram(alpha=.5, binwidth=1, position = 'identity') +
  geom_vline(xintercept=c(5,10,15))
dev.off()

################### using nanopore, filtering deepvariant calls by qual is not good.
```
